---
title: "Covid91 vaccine study Final2022"
author: "Teresa Juan"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final_Data.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
Males <- filter(FinalData, sex=="M")
```

```{r}
Males <- select(Males, -sex)
```


## Graphic Description 

```{r}
dd2 <- Males %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected, y=count,fill= treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In order to compare each sub group we toke them apart and did a separate data for each group. First we are doing the males data, We ar going to use the ggplot data. In the chart we created for the Male group we can see that the ones that had covid19 and took the vaccine were better then the ones that took the placebo, and the Males that were OK and took the placebo had a higher chance of getting sick of covid19 then those that took the vaccine. We can see that the vaccine has a positive reaction to the males.

## Numeric Summeries

```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
By looking at the numeric summery we can see that the plot was right, the males that took the placebo 70% of them got sick of covid19 and only 32% of them that had the vaccine got it. This shows us that the vaccine was very effective because only 1.28% of the males out of 100 got sick.


## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fishtestm <-fisher.test(table1)
print(fishtestm)
```

The fisher exact test gave us a low P- value of 2.645e-11 seeing this numbers means that these results are not likely to occur by a chance. so if you are a Male that took the vaccine the odd of you getting covid19 is very low as of those who took the placebo, which is 2x greater then if you did take the vaccine.

# Eficcacy

```{r}
options(digits = 1)
maleffic=fishtestm$estimate/(1+fishtestm$estimate)*100
malefficlo=fishtestm$conf.int[1]/(1+fishtestm$conf.int[1])*100
maleffichi=fishtestm$conf.int[2]/(1+fishtestm$conf.int[2])*100
print(maleffic)
print(malefficlo)
print(maleffichi)
```
As we look at how effective is the vaccine for the Males we can see that it is very effective, being that the odd ratio is very high to 68% being above50%. The confidence interval of the effectiveness of the vaccine on Males has a low of `r malefficlo` and a high of `r maleffichi`.


# Females

```{r}
Females <- filter(FinalData, sex=="F")
```

```{r}
Females <- select(Females, -sex)
```

## Graphical Description

```{r}
dd2 <- Females %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected, y=count,fill= treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Now on our data for Females we can see that the vaccine is working  about the same as for the Males. The chart shows us that females  that took the placebo got sick of covid19 about 60% more then those that took the vaccine. Females  that took the vaccine were about 40% chance of getting covid19. seeing that more females  get sick with the placebo and less with the vaccine we can conclude that the vaccine is having a positive results on females as well. And the females that are on the OK side have a 50% to 50% chance of getting covid19 whether they take the vaccine or the placebo.

## Numeric Summeries

```{r}
table2 <- xtabs(~infected + treatment, data=Females)
rowPerc(table2)
colPerc(table2)
```
In the numeric result we can see that Females that took the placebo 2.3% of them got Covid19, and the females that took the vaccine only 1.62% of then got Covid19. Just by seeing this results we can assume that the vaccine had a positive effect on females because there was low chance of getting Covid19 then if you didn't take it. Females are at 2 times at risk of getting Covid19 if they don't take the vaccine.

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fishtestf <-fisher.test(table2)
print(fishtestf)
```

The fisher exact test gave us a very low p-value 9e-04, being that is such a small number , the chances of this results are not likely to happen by chance. As we look at the ratio, we got a 1.43 odd ratio, meaning that the chances of getting Covid19 is 1.43% time likely to get it if they thake the placebo, then if they take the vaccine.

# Efficacy

```{r}
femeffic=fishtestf$estimate/(1+fishtestf$estimate)*100
femefficlo=fishtestf$conf.int[1]/(1+fishtestf$conf.int[1])*100
femeffichi=fishtestf$conf.int[2]/(1+fishtestf$conf.int[2])*100
print(femeffic)
print(femefficlo)
print(femeffichi)
```
By looking on the odd ratio we can see that the vaccine is effective for Females as well, since the ratio is above 50% being at 59% .The confidence interval of the effectiveness of the vaccine on females has a low of `r femefficlo` and a high of `r femeffichi`.



# LGBTQ

```{r}
LGBTQ <- filter(FinalData,LGBTQ =="gay")
```

## Graphical Description

```{r}
dd2 <- LGBTQ %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected, y=count,fill= treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Looking at the bar-plot  for "Gay" we notice that the vaccine is not being effective on these group. Being that the percentage  of gay that took the vaccine is really high at 85% and got sic with covid19. And 25% of them that  took the placebo got sick with covid19. Seeing this results we can see that the vaccine wasn’t working so well on Gays then on the Female and Males.

## Numeric Summeries

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table3)
colPerc(table3)
```

In the numerical result we see that 3% of the LGBTQ that took the vaccine got sick with covid19, and only 0.5% of them that took the placebo got sick. Being the number so high it meas that the vaccine was ineffective on LGBTQ.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fishtestL <-fisher.test(table3)
print(fishtestL)
```

The odd ratio is telling us that the  placebo is  0.1%. of catching vovid19, this group  is better off with the placebo then the vaccine.

# Efficacy

```{r}
Leffic=fishtestL$estimate/(1+fishtestL$estimate)*100
Lefficlo=fishtestL$conf.int[1]/(1+fishtestL$conf.int[1])*100
Leffichi=fishtestL$conf.int[2]/(1+fishtestL$conf.int[2])*100
print(Leffic)
print(Lefficlo)
print(Leffichi)
```

# Druggies
## Graphical Description 
## Numeric Summeries
## Inferential Results
# Conclusion 
